### 2016-04-04 Manuscript/BiologyLetters/response2.tex

We emphasised the importance of skeletal characters in the introduction as follows:
\textit{``A downside of this method is that it requires molecular data for living taxa and discrete morphological/anatomical data shared between both living and fossil taxa (i.e. hard tissue characters such as skeletal features).''}
lines @@@.

@@

We add the following sentence to the discussion:
\textit{``Additionally, our analyses may underestimate the problem as we do not distiguish between soft and hard tissue characters; if a living taxon has only soft tissue coded anatomical characters then it will not have overlapping data with fossils that only have hard tissues preserved.''}
lines @@@.

\item{\textcolor{blue}{I'm also concerned that my original comments about the biases in character sets included still aren't adequately addressed; they now acknowledge that they didn't include any character sets that weren't available in print, but that doesn't address what kind of difference it would make to their results. I'm concerned it could be a really important bias; morphological data was much more commonly published for extant taxa in older phylogenetic analyses, and those are less likely to be available digitally. You might be substantially underestimating the availability of these data, and if TE phylogenies aren't including them, I'm rather disappointed in the folks constructing them. }}

We agree with this comment (and Referee 3 comment 2) and modified the paragraph discussing this caveat as follows:
\textit{``Although this will bias our results towards lower coverage we do not think this bias will be large, as many recent morphological matrices reuse living taxa characters from older matrices (see ESM1), so many of these data will be present in our analyses.
Additionally, these older matrices are likely to differ from more recent ones in terms of their underlying definition of homology and their coding practices (see [19]).
Therefore, care needs to be taken when deciding how to include these older matrices.''} lines @@@.

Since we did not collect these ``older'' matrices, we are unable to properly quantify the bias introduced by ignoring them.
To test it we should rerun the whole analysis from the data collection to include them and to see the difference between the present data and the ones including these ``older'' matrices.
This would really interesting but would also take a substantial amount of time not available at present.

\item{\textcolor{blue}{I don't agree with the assertion that the genus level should be our primary concern; more and more studies are building species-level trees with fossil taxa. I'm happy for the authors to discuss the patterns at the genus level, because it is one of the important levels, but given that genera are an artificial line in the taxonomic sand, it's certainly not the primary one, and I'd hate to see this statement sail unopposed into the literature. I'd like to see it moderated a bit in the final version of the manuscript.}}

>>>>

We emphasised the importance of skeletal characters in the introduction as follows:
\textit{``A downside of this method is that it requires molecular data for living taxa and discrete morphological/anatomical data shared among both living and fossil taxa (i.e. hard tissue characters such as skeletal features).''}
lines @@@.
% NC: Add line numbers

@@

We add the following sentence to the discussion:
\textit{``Additionally, our analyses may underestimate the problem as we do not distiguish between soft and hard tissue characters; if a living taxon has only soft tissue coded anatomical characters then it will not have overlapping data with fossils that only have hard tissues preserved.''}
lines @@@.
% NC: Add line numbers
\item{\textcolor{blue}{I'm also concerned that my original comments about the biases in character sets included still aren't adequately addressed; they now acknowledge that they didn't include any character sets that weren't available in print, but that doesn't address what kind of difference it would make to their results. I'm concerned it could be a really important bias; morphological data was much more commonly published for extant taxa in older phylogenetic analyses, and those are less likely to be available digitally. You might be substantially underestimating the availability of these data, and if TE phylogenies aren't including them, I'm rather disappointed in the folks constructing them. }}

% NC: I've rewritten this to make it clearer what you have done and to stop it reading like "yeah we know there is a bias but we are too lazy to fix it". Now it says "yeah we know there is a bias but we are too lazy to fix it, but we also think it isn't a big bias and would add more bias anyway". A much better justification.
We acknowledge this comment (and Referee 3 comment 2) and have now made more explicit reference to this bias. 
However, because we did not collect these matrices, we are unable to properly quantify the bias introduced by excluding them.
To test this we would need to collect these data and then rerun the analysis.
This would really interesting but would also take a substantial amount of time to collect the data, which is not available at present. 
Additionally we do not believe this bias will be as large as the reviewer fears.
Many recent matrices reuse characters from older matrices, thus it is likely most of the data from major previous studies are present in our analyses.
We also share the concerns of Peter Wagner that using ``old'' matrices may introduce a different kind of bias into phylogenies, as definitions of homology and characters have changed over time, making the kind of matrix we are unlikely to find with our search protocol harder to incorporate into Total Evidence analyses.
We therefore modified the paragraph discussing this bias as follows:
\textit{``Although this will bias our results towards lower coverage we do not think this bias will be large, as many recent morphological matrices reuse living taxa characters from older matrices (see ESM1), so many of these data will be present in our analyses.
Additionally, these older matrices are likely to differ from more recent ones in terms of their underlying definition of homology and their coding practices (see [19]).
Therefore, care needs to be taken when deciding how to include these older matrices.''} lines @@@.
% NC: Add line numbers
\item{\textcolor{blue}{I don't agree with the assertion that the genus level should be our primary concern; more and more studies are building species-level trees with fossil taxa. I'm happy for the authors to discuss the patterns at the genus level, because it is one of the important levels, but given that genera are an artificial line in the taxonomic sand, it's certainly not the primary one, and I'd hate to see this statement sail unopposed into the literature. I'd like to see it moderated a bit in the final version of the manuscript.}}
